Materils table not populating
Radial build figure bigger - relative to legend
Radial build figure reference
rounding in Table 4
C21010101 not populating for shield
C21010102 not populating for shield
C21010103 not and 04 not
Primary structure and support is outside of the vessel - this is what allows the 1g lateral forces to be taken up for earthquake-proofing

Spacing in the vacuum system costs
Why are vacuum vessel adn cooling systems identical costs?

Power supply costs are too low
plasma gun/formation costs should be 3-10M
Assy and instillation costs C2201011 needs replacing in the text with C210111
Coolant primaryC and secondaryC need replacing
Fuel handling and storage needs to be replaced by ITER 2011 costing, possibly a reference to the work that Mike has done on this topic.

special materials needs some thought - this is where the PbLi will be put.
Contingency can be ruled out in NOAK, 438M savings

C9100000 in 32 needs to be overwritten or changed to 320000
C920000 in 36 needs to be changed.

40 - percent to \%
We did replace this with a function proportional to the construction time... need to pull it in.

52 - spare parts - 10\% of all capital costs - this should just be C22, but note that this is in the LCOE calculation section already, so it should be small... applied only to BOP

54 insurance should be 1M
55 initial fuel load - do we need Kg of tritium?  Perhaps it will be less.
C580000 is not being calculated - this is the decommissioning costs

Contingency for noak should be smaller!

C630000 not being calculated correctly

% in table 10 should be every line?
C990000 not being calculated
is the 7367 ithout the indirects?  If so, then costs will be much higher - include the same fractions as Novatron

O&M costs are too high
SCR is in LCOE sextion this should be C22 10% or so per annum C750000 is not being replaced

LCOE not being calculate correctly  

Cost elements of the LCOE claculation not being generated correctly.

Remove the NOAK discussion for now.
