\ifsafetycosts
\section*{Safety and hazard cost considerations}

This costing report includes additional cost contributions motivated by the
safety and hazard analysis work (for example Lukacs \& Williams 2020 and
related references). The following effects are explicitly reflected in the
cost accounting when safety and hazard mitigation costs are included:

\begin{itemize}
  \item \textbf{Radioactive release of tritium from storage (land and land rights)}:\\
  A tritium-driven site boundary model is used to estimate the land area
  required to keep off-site doses within emergency reference levels for
  tritium and dust releases. The associated site cost is captured as an
  additional land and land-rights component \(C110200\) in Cost Category 11,
  with the total land cost given by \(C110000 = C110100 + C110200\).

  \item \textbf{Plant licensing (region-specific)}:\\
  Region-dependent licensing and regulatory costs for fusion facilities
  are incorporated into Cost Category 13 – Plant Licensing. These costs
  account for fees and regulatory activities under the relevant national
  framework (for example NRC byproduct licensing in the US or REPPIR/EA
  requirements in the UK).

  \item \textbf{Plasma disruption mitigation system (tokamak-specific)}:\\
  For magnetic confinement designs, a disruption mitigation system based
  on shattered pellet injection is included as Cost Category 22.01.20
  (Disruption mitigation system) under Heat Island Components.

  \item \textbf{Hot-cell / remote handling equipment}:\\
  Remote handling and hot-cell equipment needed to manage activated
  components is represented as Cost Category 22.06.06 (Remote handling
  equipment) under Other Heat Island Equipment.

  \item \textbf{Third-party liability insurance}:\\
  An annual third-party liability insurance premium is estimated by
  scaling from fission experience using the Level of Safety Assurance
  (LSA) framework. The corresponding cost contribution appears in
  Cost Category 78 – Taxes and Insurance.
\end{itemize}

Unless otherwise noted, the numerical values for these contributions are
taken from the safety implementation documentation and mapped consistently
into the CAS cost structure used throughout this report.
\fi

