\subsubsection{Cost Category 22.7: Instrumentation and Control }

For a NOAK plant, it is anticipated that the I\&C technologies will be much more advanced than current  capabilities. This system will include the power core instrumentation and control, radiation and  monitoring equipment, isolated indicating and recording equipment, data acquisition and  recording, communication equipment, benchboards, panels, racks, process computers, I\&C tubing and fittings, instrumentation, and software.\\

Most fusion I\&C estimates were are based on the Starfire  estimate of \$55 M in 2009\$ (\$79 M 2023 \$) . The three ARIES algorithms of constant values have been  escalated over the years, but still remain in the \$50-\$60 M class when expressed in 2009\$ (79-86M 2023\$).  We allocate a budget of \$ C220700 M for this cost category, which consists of the following sub categories:

\begin{itemize}
\item Cost Category 22.07.01 Power Core Instrumentation and Control Equipment. Instrumentation and control equipment to monitor and manage the plasma and all  its related functional equipment contained in the Power Core (includes all equipment in Account 22).

\item Cost Category 22.07.02 Radiation Monitoring Equipment. Monitoring equipment to ensure safe levels of radiation are maintained  during routine plant operations, determining the tritium inventories in all plant components, and  analyze/report any fault or accidental release of radioactive materials. 

\item Cost Category 22.07.03 Isolated Indicating and Recording Equipment. This account includes all  indicating and recording equipment that is physically isolated from the main power core, yet is  necessary to monitor the performance and health of the power core and its related equipment. 

\item Cost Category 22.07.04 Data Acquisition and Recording Equipment. Enormous amounts of data will need acquired and recorded  for analysis to properly control the fusion plasma and the plant. The access rate for the data  acquisition will be very high to be able to adequately control the plasma real time. This means vast data storage will be required, with the most up-to-date hardware.

\item Cost Category 22.07.05 Communications Equipment. This communication equipment will handle  all forms of communication both inter-and intra-plant. 

\end{itemize}